---
title: "Handout13Code_RMarkdown_STUDENTS"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Example 1: DeliveryTime.xls

Import Data

```{r}
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
library(readxl)  #install.packages('readxl) if you haven't already
DeliveryData=read_excel("DeliveryTime.xls")
str(DeliveryData)
```

### 1(a)

Original model

```{r}
fit.dlvr=lm(Deliverytime ~ NumberCases + Distance, data = DeliveryData)
summary(fit.dlvr)
```

### 1(b-e)

Standardized Variable model.\
Use the function scale() for every variable in the model.

```{r}
fit.dlvr.std=lm(scale(Deliverytime)~scale(NumberCases) + scale(Distance), data=DeliveryData)
summary(fit.dlvr.std)
```

### 1(f)

```{r}
x0=data.frame(NumberCases=10,Distance=200)
predict(fit.dlvr,x0)
predict(fit.dlvr.std, x0)


```

# Example 2: ArticSeaIce.xls

Import Data

```{r}
#setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
#library(readxl) 
ArticData=read_excel("ArticSeaIce.xls")
str(ArticData)

```

### 2(a)

Linear model

```{r}
fit.artic1=lm(Year ~ SASIE, data = ArticData)
summary(fit.artic1)

par(mfrow=c(2,2))
plot(Year ~ SASIE,data=ArticData, pch=19)
abline(fit.artic1,col='blue')
plot(resid(fit.artic1)~fitted(fit.artic1), xlab='Fitted', ylab='Residuals');abline(h=0,lty=2)
qqnorm(resid(fit.artic1),ylab='Residuals'); qqline(resid(fit.artic1),lty=2)
hist(resid(fit.artic1),main="", xlab="Residuals",col='lightblue')
par(mfrow=c(1,1))
```

### 2(b)

######################################## 

# Illustration: rate of dec change

######################################## 

```{r}
plot(SASIE~Year,data=ArticData, pch=19,col='red')

fit.artic2002=lm(SASIE[Year<=2002]~Year[Year<=2002], data=ArticData)
summary(fit.artic2002)
abline(fit.artic2002,col=1)

fit.artic2004=lm(SASIE[Year<=2004]~Year[Year<=2004], data=ArticData)
summary(fit.artic2004)
abline(fit.artic2004,col=2)

fit.artic2006=lm(SASIE[Year<=2006]~Year[Year<=2006], data=ArticData)
summary(fit.artic2006)
abline(fit.artic2006,col=3)

fit.artic2008=lm(SASIE[Year<=2008]~Year[Year<=2008], data=ArticData)
summary(fit.artic2008)
abline(fit.artic2008,col=4)

fit.artic2010=lm(SASIE[Year<=2010]~Year[Year<=2010], data=ArticData)
summary(fit.artic2010)
abline(fit.artic2010,col=5)

fit.artic2012=lm(SASIE[Year<=2012]~Year[Year<=2012], data=ArticData)
summary(fit.artic2012)
abline(fit.artic2012,col=6)

legend('bottomleft',col=1:6,lty=1,
       legend=c('2002','2004','2006','2008','2010','2012'))
```

######################################## 

# End of Illustration

######################################## 

### 2(c)

Quadratic model

```{r}
fit.quad=lm(SASIE~Year+I(Year^2),data=ArticData)
summary(fit.quad)
```

Plots

```{r}
#Plot regression curve
coeff=fit.quad$coeff
par(mfrow=c(1,4))
plot(SASIE~Year,data=ArticData,pch=19)
curve(coeff[1]+coeff[2]*x+coeff[3]*x^2,xlim=range(ArticData$Year),add=T,col='blue')

#Residual plots
plot(resid(fit.quad)~fitted(fit.quad), xlab='Fitted', ylab='Residuals');abline(h=0,lty=2)
qqnorm(resid(fit.quad),ylab='Residuals'); qqline(resid(fit.quad),lty=2)
hist(resid(fit.quad),main="", xlab="Residuals",col='lightblue')
```

VIF

```{r}
library('car')
vif(fit.quad)
```

### 2(d)

Fit the regression model centering the year variable.

```{r}
ArticData$YearC=ArticData$Year-mean(ArticData$Year)
fit.quadC=lm(SASIE~YearC+I(YearC^2),data=ArticData)
summary(fit.quadC)

#Plot regression curve
coeffC=fit.quadC$coeff
plot(SASIE~Year,data=ArticData,pch=19)
curve(coeffC[1]+coeffC[2]*x+coeffC[3]*x^2,xlim=range(ArticData$Year),add=T,col='blue')

#Residual plots
plot(resid(fit.quadC)~fitted(fit.quadC), xlab='Fitted', ylab='Residuals');abline(h=0,lty=2)
qqnorm(resid(fit.quadC),ylab='Residuals'); qqline(resid(fit.quadC),lty=2)
hist(resid(fit.quadC),main="", xlab="Residuals",col='lightblue')
```

```{r}
vif(fit.quadC)
```

```{r}
mean(ArticData$Year)
```

# Example 3

```{r}
cakes=read_excel("cakes.xlsx")
str(cakes)
```

```{r}
fit.cake=lm(Palatability ~ Time + Temperature, data = cakes)
summary(fit.cake)

```

Examine the residuals vs fitted ad residuals vs each predictor

```{r}
par(mfrow=c(2,2))

```
