---
title: "Handout1_RMarkdown"
output: html_document
---

# Example 1: airfare

Import Data Into R from Excel file\
(1) Run command: install.packages("readxl") if have never done so\
(2) Save data file to same location as R script; then,\
(3) Run the following three commands

```{r}
library(readxl)
setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) 
airfareData=read_excel("airfare.xls")
```

View data structure and summary\
syntax str(my.data), summary(my.data)

```{r}
str(airfareData)
summary(airfareData)
```

### 1(f)

Create scatter plot\
syntax: plot(y\~x, data=my.data)

```{r}
plot(price~distance,data=airfareData) 
```

The pch argument changes the symbol of points\
pch=19 makes the points solid circles

```{r}
plot(price~distance,data=airfareData, pch=19) 

```

# Example 3: carsMSRP

Import Data Into R from .csv\
(1) Run command: install.packages("readr") if have never done so\
(2) Save data file to same location as R script; then,\
(3) Run the following three commands

```{r}
library(readr)
setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) 
carsMSRP = read_csv("carsMSRP.csv")
```

Correlation matrix

```{r}
cor(carsMSRP[carsMSRP$Year==2009 & carsMSRP$Popularity<2400
             & carsMSRP$HP<450 & carsMSRP$City_MPG<30,c(13,14,5,15)])

```
