---
title: "Handout4Code_RMarkdown_students"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Example 1: MaleBirths.xls

Import Data Into R from Excel file\
(1) Run command: install.packages("readxl") if have never done so\
(2) Save data file to same location as R script; then,\
(3) Run the following three commands

```{r}
library(readxl)
setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) 
MaleBirths=read_excel("MaleBirths.xls")

```

View data frame structure -*-*-*-* IMPORTANT: Make sure you substitute "???" by the correct code.-*-*-*-*

```{r}
str(???)
View(???)
```

### 1(a & (e))

Fitted line plots for each country

```{r}
par(mfrow=c(1,3)) #1x3 plot grid
```

USA

```{r}
plot(usa~year, data=MaleBirths, ylab='male birth percentage',pch=19,ylim=c(50.8,51.6),main='USA')
fit_usa=lm(usa~year, data=MaleBirths)
abline(fit_usa,col='blue')
```

Canada

```{r}
plot(canada~year,data=MaleBirths,ylab='male birth percentage', pch=19,ylim=c(50.8,51.6),main='Canada')
fit_canada=lm(canada~year, data=MaleBirths)
abline(fit_canada,col='blue')
```

Netherlands

\-*-*-*-* IMPORTANT: Make sure you substitute "???" by the correct code.-*-*-*-*

```{r}
plot(netherla~year,data=MaleBirths,ylab='???',pch=19,ylim=c(50.8,51.6),main='???')
fit_netherla=lm(netherla~year, data=MaleBirths)
abline(fit_netherla,col='blue')
```

### 1(f-h)

Model summary % male birth regressed on year each country

\-*-*-*-* IMPORTANT: Make sure you substitute "???" by the correct code.-*-*-*-*

```{r}
summary(fit_usa)
summary(fit_canada)
summary(fit_netherla)

```

s each country

```{r}
c(summary(fit_usa)$sigma,summary(fit_canada)$sigma,summary(fit_netherla)$sigma)

```

R.squared each country

\-*-*-*-* IMPORTANT: Make sure you substitute "???" by the correct code.-*-*-*-*

```{r}
c(summary(fit_usa)$r.squared,
  summary(fit_canada)$r.squared,
  summary(fit_netherla)$r.squared)

```
