---
title: "Handout10Code_RMarkdown_STUDENTS-ANSWERS"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Example 1: NBA18-19.txt

Import Data

```{r}
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
NBAData0 = read.table('NBA18-19.txt', header=TRUE, sep="")
str(NBAData0)
```

Subset data to include point guards only

```{r}
NBAData=subset(NBAData0, PS=='PG')
head(NBAData)
```

### 1(a)

Histogram + Boxplots + Outliers

```{r}
par(mfrow=c(2,1))
with(NBAData, hist(PTS,col='lightblue', main = ""))
with(NBAData, boxplot(PTS,col='lightblue', horizontal = TRUE, xlab = "PTS"))
par(mfrow=c(1,1))

```

Those players with PTS\>2000

```{r}
cbind(NBAData$Player[NBAData$PTS>2000],
      NBAData$PTS[NBAData$PTS>2000])

```

### 1(c-d)

Create variables eff48m and MPG

```{r}
NBAData$eff48m=(NBAData$PTS + NBAData$TR + NBAData$AS + NBAData$ST + 
        NBAData$BK - (NBAData$FGA-NBAData$FGM) - 
        (NBAData$FTA-NBAData$FTM) - NBAData$TO)*48/NBAData$Min
NBAData$MPG=NBAData$Min/NBAData$GP
head(NBAData)
```

Model

IMPORTANT: Do not forget to delete the "\#" so you can actually run the code.

```{r}
NBA.fit0=lm(eff48m ~ MPG, data = NBAData)
summary(NBA.fit0)



```

Fitted line plot eff48m vs MPG

```{r}
par(mfrow=c(1,3))
plot(eff48m~MPG, pch=19,data=NBAData)
abline(NBA.fit0,col='blue')
plot(NBA.fit0,c(1,2))
```

### 1(e-f)

```{r}
par(mfrow=c(1,1))
plot(eff48m~MPG, pch=19,data=NBAData)
abline(NBA.fit0,col='blue')
#text labels
text(NBAData$MPG[c(33,45,83,103,108)], NBAData$eff48m[c(33,45,83,103,108)],
     label=NBAData$Player[c(33,45,83,103,108)],pos=c(4,2,4,4,2),cex=0.9)
```

```{r}
NBAData[NBAData$Player == "ulis,tyler",]

```

### 1(g)

Subset data to include only full time point guards

```{r}
NBAFull=subset(NBAData,Min>500 & MPG>6.09)
str(NBAFull)

```

### 1(h1)

Model

IMPORTANT: Do not forget to delete the "\#" so you can actually run the code.

```{r}
NBA.fit=lm(eff48m ~ MPG, data = NBAFull)
summary(NBA.fit)


```

Fitted line plot eff48m vs MPG & Residual Diagnostics

```{r}
par(mfrow=c(1,3))
plot(eff48m~MPG, data=NBAFull,pch=19)
abline(NBA.fit,col='blue')
id=c(19, 31, 37, 43, 79)
text(NBAFull$MPG[id],NBAFull$eff48m[id],label=NBAFull$Player[id],pos=c(4,2,2,2,2),cex=0.6)
plot(NBA.fit, c(1,2))
par(mfrow=c(1,1))
```

### 1(h2)

Note: we chose MPG=24.7 b/c its the mean MPG

```{r}
mean(NBAFull$MPG)
```

Predict efficiency of player with avg. MPG (ie. MPG=24.7)

(1) Algebraic

```{r}
NBA.fit$coefficients[1] + NBA.fit$coefficients[2]*24.7


```

(2) Matrix\
    tip1: use matrix multiplication.\
    tip2: use your fitted model called NBA.fit and get the coefficients from this model using "NBA.fit\$XXXX".

IMPORTANT: Do not forget to delete the "\#" so you can actually run the code.

```{r}
c(1,24.7)%*%c(NBA.fit$coefficients[1],NBA.fit$coefficients[2])


```

(3) Ysing predict()

IMPORTANT: Do not forget to delete the "\#" so you can actually run the code.

```{r}
predict(NBA.fit, data.frame(MPG=24.7))


```

Suppose we followed all point guards who play 24.7 minutes per game. Predict the average efficiency for all these players (one value).

```{r}
mean(predict(NBA.fit, data.frame(MPG=24.7)))
```

### 1(k)

View Westbrook's stats

```{r}
NBAFull[NBAFull$Player=="westbrook,russel",]

```

Predict efficiency of player with MPG=36.1 (Westbrook's MPG) using algebra.

```{r}

NBA.fit$coefficients[1] + NBA.fit$coefficients[2]*36.1


```

Predict efficiency of player with MPG=36.1 (Westbrook's MPG) using predict()

```{r}
c(1,36.1)%*%c(NBA.fit$coefficients[1],NBA.fit$coefficients[2])

predict(NBA.fit, data.frame(MPG=36.1))


```

### 1(l)

Compute SE's for CI/PI

IMPORTANT: Do not forget to delete the "\#" so you can actually run the code.

```{r}
X=matrix(c(rep(1,82),NBAFull$MPG),nrow=82)
X
x0=c(1,36.1)
x0
s=summary(NBA.fit)$sigma
se_p=s*sqrt(1+ t(x0)%*%solve(t(X)%*%X)%*%x0)
se_c=s*sqrt(t(x0)%*%solve(t(X)%*%X)%*%x0)
se_p
se_c
```

### 1(m)

CI and PI for efficiency for MPG=36.1 (Westbrook's MPG)

```{r}
#(tstar=qt(p=???, df=???))
#????
ts=qt(p=0.975, df=80)
ts

```

### 1(n)

CI and PI for efficiency for MPG=36.1 (Westbrook's MPG) using predict()

```{r}
new_MPG=data.frame(MPG=36.1)
predict(NBA.fit, new_MPG, interval='confidence', level = .95)
predict(NBA.fit, new_MPG, interval='prediction', level = .95)

```

### 1(o2)

"Pointwise" Confidence/Prediction bands

```{r}
#Harden's stats
NBAFull[NBAFull$Player=="harden,james",]
```

```{r}
par(mfrow=c(1,1))
plot(eff48m~MPG, data=NBAFull,pch=19)
abline(NBA.fit,col='blue')
text(NBAFull$MPG[id],NBAFull$eff48m[id],label=NBAFull$Player[id],pos=c(4,2,2,2,2),cex=0.6)

library(RColorBrewer); cols <- brewer.pal(2, "Dark2")

#Plot "pointwise" Confidence/Prediction bands 
new_MPG=data.frame(MPG=seq(10,40,by=1))
CI.Limits=as.data.frame(predict(NBA.fit,new_MPG,interval='confidence',level=0.95))
PI.Limits=as.data.frame(predict(NBA.fit,new_MPG,interval='prediction',level=0.95))

lines(new_MPG$MPG, CI.Limits$lwr,col=cols[1])
lines(new_MPG$MPG, CI.Limits$upr,col=cols[1])

lines(new_MPG$MPG, PI.Limits$lwr,col=cols[2])
lines(new_MPG$MPG, PI.Limits$upr,col=cols[2])

legend('topleft',legend=c('CI','PI'),lty=1,col=cols)
```

# Example 2: houses.txt

Import data

```{r}
houses = read.table('houses.txt', header=TRUE, sep="")
str(houses)
plot(Price~Size, data=houses)
```

Scatter plot log(y)\~x

```{r}
plot(log(Price,10)~Size, data=houses)

```

Residual plots for log-linear model

```{r}
par(mfrow=c(1,2))
plot(lm(log(Price,10)~Size, data=houses),c(1,2))
par(mfrow=c(1,1))
```

Fit log-linear model

```{r}
fit.houses=lm(log(Price,10)~Size,data=houses)
summary(fit.houses)

```

### 2(a)

PI 2000 ft\^2 house

```{r}
new_houses=data.frame(Size=2000)
predict(fit.houses, new_houses, interval='prediction', level = .95)

```

Back-tranformed PI 2000 ft\^2 house 

```{r}
exp(predict(fit.houses, new_houses, interval='prediction', level = .95))



```

# Example 3: Cars75

Import Data

```{r}
Cars75Data=read.table("cars75.txt",header=T, sep='\t')
str(Cars75Data)

```

### 3(a)

Fit 1-variable model with horsepower

```{r}
cars.fit=lm(mpg~horsepower,data=Cars75Data)
summary(cars.fit)


```

CI/PI horsepower=130

```{r}
new_cars=data.frame(horsepower=130)
predict(cars.fit, new_cars, interval='confidence', level = .95)
predict(cars.fit, new_cars, interval='prediction', level = .95)
```

### 3(b)

Fit 2-variable model with horsepower & width

```{r}
cars.fit2=lm(mpg~horsepower+width,data=Cars75Data)
summary(cars.fit2)

```

CI/PI horsepower=130, width=70

```{r}
new_cars2=data.frame(horsepower=130, width=70)
predict(cars.fit2, new_cars2, interval='confidence', level = .95)
predict(cars.fit2, new_cars2, interval='prediction', level = .95)

```

### 3(c)

CI/PI horsepower=130, width=65

```{r}

new_cars3=data.frame(horsepower=130, width=65)
predict(cars.fit2, new_cars3, interval='confidence', level = .95)
predict(cars.fit2, new_cars3, interval='prediction', level = .95)

```
