---
title: "Handout6Code_RMarkdown_studentsANSWERS"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Example 2: TVs and Life Expectancy

Import Data Into R from Excel file\
(1) Run command: install.packages("readxl") if have never done so\
(2) Save data file to same location as R script; then,\
(3) Run the following three commands

```{r}
library(readxl)
setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) 
TVLife=read_excel("TVLife.xls")

```

### 2(a)

\-*-*-*-* IMPORTANT: Make sure you substitute "???" by the correct code.-*-*-*-*

Fit linear model and Fitted line plot

```{r}
par(mfrow=c(1,1)) #plot in 1x1 grid
plot(LifeExpectancy ~ TVsperK, data=TVLife, pch=19, col='black')
tv.fit=lm(LifeExpectancy ~ TVsperK, data=TVLife)
abline(tv.fit, col='blue')

```

### 2(b)

Identify points.\
Add text labels to points of interest

```{r}
par(mfrow=c(1,1)) #plot in 1x1 grid
plot(LifeExpectancy~TVsperK, data=TVLife, pch=19,col='black')
tv.fit=lm(LifeExpectancy~TVsperK, data=TVLife)
abline(tv.fit$coeff,col='blue')

with(TVLife, text(x=TVsperK[c(15,16,22,20)],y=LifeExpectancy[c(15,16,22,20)],
                  labels=c(country[c(15,16,22)],'US'), pos=c(4,4,4,1), cex=0.8))

```

### 2(d)

\-*-*-*-* IMPORTANT: Make sure you substitute "???" by the correct code.-*-*-*-*

PLOT: residuals vs. fitted plot

```{r}
plot(resid(tv.fit)~fitted(tv.fit),ylab="Residual",xlab="Fitted",pch=19)
abline(h=0,lty=2)

```

### 2(e)

\-*-*-*-* IMPORTANT: Make sure you substitute "???" by the correct code.-*-*-*-*\
PLOT: residuals vs. explanatory plot

```{r}
plot(resid(tv.fit) ~ TVsperK , data=TVLife, ylab="Residual", pch=19)
abline(h=0,lty=2)

```

### 2(f)

PLOT: normal Q-Q plot and histogram of residuals

```{r}
par(mfrow=c(1,2))  #plot in 1x2 grid   
qqnorm(resid(tv.fit),ylab="Residual"); qqline(resid(tv.fit),lty=3)
hist(resid(tv.fit),main="", xlab="Residual",col='lightblue')

```

### 2(g)

Statistical tests of regression assumptions

```{r}
#Shapiro-Wilk test
shapiro.test(resid(tv.fit))

#Breusch-Pagan
#install.packages('lmtest') if haven't already
library('lmtest')
bptest(tv.fit)

```

### 2(h)

View R's four-in-one diagnostic plot (will cover the right two plots later)

```{r}
par(mfrow=c(1,4)); plot(tv.fit)

```
