---
title: "Handout7Code_RMarkdown_PROF"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Example 1: mazda.txt

Import Data Into R from .txt file\
(1) Save data file to same location as R script; then,\
(2) Run the following commands

```{r}
setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) 
mazdaData = read.table('mazda.txt', header=TRUE, sep="\t")
str(mazdaData)
summary(mazdaData)
head(mazdaData)
```

### 1(a)

+0.5 also helps deal with log(age=0)

```{r}
mazdaData$Age=91-mazdaData$Year+0.5  
head(mazdaData)
summary(mazdaData)
```

### 1(b)

Price\~Age

```{r}
plot(Price~Age,data=mazdaData,pch=19)
mazda.fit0=lm(Price~Age,data=mazdaData)
abline(mazda.fit0,col='blue')
summary(mazda.fit0)
```

### 1(c)

Residual diagnostics\
Plots to examine whether assumptions are met.

```{r}
par(mfrow=c(2,2))
plot(Price~Age,data=mazdaData,pch=19)
mazda.fit0=lm(Price~Age,data=mazdaData)
abline(mazda.fit0,col='blue')
summary(mazda.fit0)
plot(mazda.fit0,c(1,2))
hist(resid(mazda.fit0),main="", xlab="Residuals",col='lightblue')
par(mfrow=c(1,1))

```

### 1(e)

Transforming X only.\
Price\~log(Age,2)

```{r}
mazda.fit1=lm(Price~log(Age,2),data=mazdaData)
summary(mazda.fit1)

```

Residual diagnostics\
Plots to examine whether assumptions are met.

```{r}
par(mfrow=c(2,2))
plot(Price~log(Age,2),data=mazdaData,pch=19, xlab='log(Age)')
abline(mazda.fit1,col='blue')
plot(mazda.fit1,c(1,2))
hist(resid(mazda.fit1),main="", xlab="Residuals",col='lightblue')
```

### 1(g)

Transforming Y only.\
log(Price,2)\~Age

```{r}
mazda.fit2=lm(log(Price,2)~Age,data=mazdaData)
summary(mazda.fit2)

```

Residual diagnostics\
Plots to examine whether assumptions are met.

```{r}
par(mfrow=c(2,2))
plot(log(Price,2)~Age,data=mazdaData, pch=19, ylab='log(Price)')

abline(mazda.fit2,col='blue')
plot(mazda.fit2,c(1,2))

hist(resid(mazda.fit2),main="", xlab="Residuals",col='lightblue')
par(mfrow=c(1,1))
```

### 1(i)

log(Price,2)\~log(Age,2)

```{r}
mazda.fit3=lm(log(Price,2)~log(Age,2),data=mazdaData)
summary(mazda.fit3)

```

### 1(j)

Box-Cox Transformation

```{r}
#install.packages('car') #if havn't already
library(car)
powerTransform(Price~Age,data=mazdaData)
```

# Example 2: AnimalBrains.xls

Import Data Into R from Excel file\
(1) Run command: install.packages("readxl") if have never done so\
(2) Save data file to same location as R script; then,\
(3) Run the following three commands

```{r}
library(readxl)
#setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) 
AnimalBrains=read_excel('AnimalBrains.xls')
```

View Data Structure

```{r}
str(AnimalBrains)
head(AnimalBrains)
summary(AnimalBrains)
```

### 2(a)

Linear model (body \~ geration)

```{r}

  

```

Original untransformed.\
Plots to examine whether assumptions are met.

```{r}



```

Identify transformations.\
Start by transforming Y going down the ladder, then try transforming X.

```{r}

```

```{r}

```

```{r}

```

```{r}

```

### 2(f)

Matrix scatter (quantitative variables only)

```{r}
pairs(AnimalBrains[,c(3,2,4,5)]) 
```

Correlation matrix

```{r}
cor(AnimalBrains[,c(3,2,4,5)])
```

### 2(g)

Matrix scatter & correlation matrix LOGGED variables.Tip: use the "log" function".

```{r}




```

### 2(h)

Model: log(body)\~log(litter)+log(gestation)

```{r}




```

3d scatterplot log(brain)\~log(body)+log(gestation)

```{r}
fit.body=lm(log(body)~log(litter)+log(gestation),data=AnimalBrains)
summary(fit.body)
library(rgl)
with(AnimalBrains, plot3d(x=log(litter), y=log(gestation), z=log(body), size=10))

#regression surface function
surface.body<-function(x1,x2){
  fit.body$coeff[1]+fit.body$coeff[2]*x1+fit.body$coeff[3]*x2
}

#plot regression surface
x1<-seq(min(log(AnimalBrains$litter)),max(log(AnimalBrains$litter)),length=15)
x2<-seq(min(log(AnimalBrains$gestation)),max(log(AnimalBrains$gestation)),length=15)
y<-outer(x1,x2,surface.body)
persp3d(x1,x2,y,alpha=0.2,col='lightblue',add=TRUE)
persp3d(x1,x2,y,alpha=1,col='black',front = "lines", back = "lines",add=TRUE)
```

### 2(i)

Residual diagnostics\
Plots to examine whether assumptions are met.

```{r}
par(mfrow=c(2,3)) 
plot(resid(fit.body)~fitted(fit.body),ylab="Residuals",xlab="Fitted",pch=19)
abline(h=0,lty=2)

#residuals vs. explanatory plot
plot(resid(fit.body)~log(litter),data=AnimalBrains,ylab="Residuals",pch=19)
abline(h=0,lty=2)

plot(resid(fit.body)~log(gestation),data=AnimalBrains,ylab="Residuals",pch=19)
abline(h=0,lty=2)

#normal Q-Q plot and histogram of residuals
qqnorm(resid(fit.body),ylab="Residuals"); qqline(resid(fit.body),lty=3)
hist(resid(fit.body),main="", xlab="Residuals",col='lightblue')

```

Shapiro-Wilk & Breusch-Pagant test

```{r}
#Shapiro-Wilk test
shapiro.test(resid(fit.body))

#install.packages('lmtest') if haven't already
library('lmtest')
bptest(fit.body)
```

# Example 3: Don't Trust Scat. Matrix

Create, x1, x2, and y

```{r}
set.seed(452) 
x1=runif(50,-4,4)
x2=-x1^2-0.5*runif(50,-4,4)^2
y=2+2*x1+0.5*x2+rnorm(length(x1),0,1.4)
set.seed(NULL) 
```

Fit model

```{r}
fit=lm(y~x1+x2)
summary(fit)
pairs(y~x1+x2)
cor(x1,x2)
```

3d scatterplot

```{r}
#library(rgl)
plot3d(x1, x2, y, size=10)

#regression surface function
surface<-function(x1,x2){fit$coeff[1]+fit$coeff[2]*x1+fit$coeff[3]*x2}

#plot regression surface
x1.surf=seq(min(x1),max(x1),length=20)
x2.surf<-seq(min(x2),max(x2),length=2)
y.surf=outer(x1.surf,x2.surf,surface)
persp3d(x1.surf,x2.surf,y.surf,alpha=0.2,col='lightblue',add=TRUE)
persp3d(x1.surf,x2.surf,y.surf,alpha=1,col='black',front = "lines", back = "lines",add=TRUE)
```

Residual diagnostics

```{r}
par(mfrow=c(2,3)) 
plot(resid(fit)~fitted(fit),ylab="Residuals",xlab="Fitted",pch=19)
abline(h=0,lty=2)

#residuals vs. explanatory plots
plot(resid(fit)~x1,ylab="Residuals",pch=19)
abline(h=0,lty=2)

plot(resid(fit)~x2,ylab="Residuals",pch=19)
abline(h=0,lty=2)

#normal Q-Q plot and histogram of residuals
qqnorm(resid(fit),ylab="Residuals"); qqline(resid(fit),lty=2)
hist(resid(fit),main="", xlab="Residuals",col='lightblue')


```

Shapiro-Wilk & Breusch-Pagant tests

```{r}
#Shapiro-Wilk test
shapiro.test(resid(fit))

#install.packages('lmtest') if haven't already
#library('lmtest')
bptest(fit)
```
